---
title: "Descriptives Practice Problems"
author: ""
date: ""
output: pdf_document
always_allow_html: true
header-includes:
  \usepackage{float}
params:
  desc_ss: NA
  desc_ans: NA
  desc_seed: NA
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, error=T, warning=F)
options(kableExtra.auto_format = FALSE)
options(knitr.table.format = "latex")
```

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

# Key
```{r echo=F, warning=F, message=F}
library(tidyverse)
library(kableExtra)
library(randomNames)
```

A research is conducting a study on IQ. They are interested in calculating the descriptive statistics of 1st year psychology students. They complete cognitive assessments on 10 psychology students and obtain the following scores:

```{r echo=F, warning=F}
sample_size <- params$desc_ss
setseed <- params$desc_seed
set.seed(setseed)
dat1 <- data.frame(Person = randomNames(n=sample_size-3, which.names = 'first', sample.with.replacement = F),
                  IQ=round(rnorm(sample_size-3, 100, 15), 0))

mode1 <- dat1[1,2]
mode2 <- dat1[3,2]
dat_mode <- data.frame(Person = randomNames(n=3, which.names = 'first', sample.with.replacement = F),
                       IQ=c(mode1, mode1, mode2))
                       
dat <- rbind(dat1, dat_mode) %>% 
  arrange(Person)
                       
kable(dat) %>% 
  kable_minimal(full_width=F) %>% 
  kable_styling(latex_options = "HOLD_position")
```

\newpage 

Help them calculate the following:

## Descriptive Statistics


### Mean


The sum of the individual scores is `r sum(dat$IQ)` and there are `r nrow(dat)` scores. Therefore, the mean of the data is `r mean(dat$IQ)`.

### Median

The scores are:

`r dat$IQ`

and when you put them in order they are:

`r sort(dat$IQ)`

Median representing the value of the middle score. If even number of scores, statistical mean of the two middle scores. The median for these scores is `r median(dat$IQ)`.

### Mode

```{r}
getMode <- function(x) {
keys <- unique(x)
keys[which.max(tabulate(match(x, keys)))]
}

mode <- getMode(dat$IQ)
```

The most frequent value of our data is `r mode`.

### Variance


```{r}
deviations <- (dat$IQ-mean(dat$IQ))
```

The deviations of our scores from the mean are:

`r deviations`

```{r}
squared_deviations <- (dat$IQ-mean(dat$IQ))^2
```

And the squared deviations are:

`r squared_deviations`

```{r}
ss <- sum(squared_deviations)
```

The sum of squared deviations is `r ss`. We can divide by $n-1$ (`r nrow(dat)`-1 = `r nrow(dat)-1`) to get the sample variance of `r round(var(dat$IQ), 2)`. 

### Standard deviation

The standard deviation is the squared `r round(sd(dat$IQ),2)`

\newpage 

# Student

A research is conducting a study on IQ. They are interested in calculating the descriptive statistics of 1st year psychology students. They complete cognitive assessments on 10 psychology students and obtain the following scores:

```{r echo=F, warning=F}
set.seed(setseed)
dat2 <- data.frame(Person = randomNames(n=sample_size-3, which.names = 'first', sample.with.replacement = F),
                  IQ=round(rnorm(sample_size-3, 100, 15), 0))

mode21 <- dat2[1,2]
mode22 <- dat2[3,2]
dat_mode2 <- data.frame(Person = randomNames(n=3, which.names = 'first', sample.with.replacement = F),
                       IQ=c(mode21, mode21, mode22))
                       
dat22 <- rbind(dat2, dat_mode2) %>% 
  arrange(Person)
                       
kable(dat22) %>% 
  kable_minimal(full_width=F) %>% 
  kable_styling(latex_options = "HOLD_position")
```

\newpage 

Help them calculate the following:

1. Mean
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline

2. Median
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
3. Mode

\newpage

4. Variance
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
5. Standard Deviation

\newpage

# Formulas

Mean = $\frac{\sum{x}}{n}$

Population variance = $\sigma^2 = \frac{\displaystyle\sum_{i=1}^{n}(x_i - \mu)^2} {n}$

Sample variance = $var = \frac{\displaystyle\sum_{i=1}^{n}(x_i - \mu)^2} {n-1}$

Standard deviation = $SD = \sqrt{\frac{\displaystyle\sum_{i=1}^{n}(x_i - \mu)^2} {n-1}}$