---
title: "Correlation Practice Problems"
author: ""
date: ""
output: pdf_document
always_allow_html: true
params:
  cor_ss: NA
  cor_ans: NA
  cor_seed: NA

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, error=T, warning=F)
options(kableExtra.auto_format = FALSE)
options(knitr.table.format = "latex")
```

```{r echo=F, warning=F, message=F}
library(tidyverse)
library(randomNames)
library(kableExtra)
seed <- params$cor_seed

sample_size <- params$cor_ss

cor_ans <- params$cor_ans
rho <- runif(1, -1, 1)
```
### Seed used is: `r seed`

# Key

A researcher is conducting a study on functioning in children. Specifically, they believes that a child's cognitive abilities (Intelligence Quotient; IQ) are associated with their emotional abilities (Emotional Quotient; EQ). To test this hypothesis, they conduct a study using `r sample_size` children. Children complete the Wechsler Intelligence Scale for Children (WISC), a measure of IQ, and the Trait Emotional Intelligence Questionnaire (TEIQue), a measure of EQ. They are interested in calculating the descriptive statistics of 1st year psychology students. They complete cognitive assessments on 10 psychology students and obtain the following scores:

```{r echo=F, warning=F, message=F}
set.seed(seed)
library(faux)
dat_cor <- rnorm_multi(n=sample_size,
                       var=2,
                       varnames = c("IQ", "EQ"),
                       mu=c(100, 50),
                       sd=c(15, 10),
                       r=rho)
dat1 <- data.frame(Person = randomNames(n=sample_size, which.names = 'first', sample.with.replacement = F))

df <- cbind(dat1, dat_cor) %>% 
  mutate(IQ=round(IQ, 0), 
         EQ=round(EQ, 0))

                       
kable(df) %>% 
   kable_styling(font_size = 15,
                 latex_options="HOLD_position") %>% 
  row_spec(row=0, bold=T)
```

\newpage 

1. What is the mean of the IQ and EQ for these individuals?

```{r}
mean_iq <- round(mean(df$IQ), 2)
mean_eq <- round(mean(df$EQ), 2)
text <- paste("The mean of IQ is ", mean_iq, " and the mean of EQ is ", mean_eq, ".", sep="")
```

```{r results="asis", eval=cor_ans=="Yes"}
cat(text)
```

2. Calculate $x*y$, $x^2$, and $y^2$. 

```{r}
df <- df %>% 
  mutate(x_y=IQ*EQ, 
         x_2 = IQ^2,
         y_2 = EQ^2)
```

```{r eval=cor_ans=="Yes"}

kable(df) %>% 
   kable_styling(font_size = 15,
                 latex_options="HOLD_position") %>% 
  row_spec(row=0, bold=T)

```


3. What is the standard deviation of IQ and EQ, and sum of the products of IQ X EQ?


```{r echo=F}
ss_iq <- round(sum((df$IQ-mean(df$IQ))^2), 2)
ss_eq <- round(sum((df$EQ-mean(df$EQ))^2), 2)


sp_iqeq <- round(sum((df$IQ-mean(df$IQ))*(df$EQ-mean(df$EQ))), 2)

text2 <- paste("The sum of squares for IQ is ", ss_iq, " and the sum of squares of EQ is ", ss_eq, ". The sum of the sproducts is ", sp_iqeq, ".", sep="")
```

```{r eval=cor_ans=="Yes", results="asis"}
cat(text2)
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

4. What is the correlation between IQ and EQ



```{r eval=cor_ans=="Yes"}
cor.test(df$IQ, df$EQ)
```


```{r eval=cor_ans=="Yes"}
text_hand <- paste("Or by manual calculations, r=SP/(sx*sy) = ",sp_iqeq, "/", "square_root(", ss_iq, " * ", ss_eq, ") = ", round(sp_iqeq/(sqrt(ss_iq*ss_eq)), 4), sep="")
```

```{r results='asis', eval=cor_ans=="Yes"}
cat(text_hand)
```


5. Create a beautiful scatterplot including line of best fit.

```{r warning=F, message=F, error=F, eval=cor_ans=="Yes" }
knitr::opts_chunk$set(fig.width=6, fig.height=5) 
library(ggplot2)
ggplot(df(), aes(IQ, EQ))+
  geom_point(color="slateblue", alpha=.9, shape=21, size=2)+
  geom_smooth(method="lm", se=F, color="black", linetype=2)+
  theme_minimal(12)
```


5. Write an results section. 

* Note, this application will not format it's answers according to APA. You must double-check that on your own. 

```{r}
p_value <- ifelse(cor.test(df$IQ, df$EQ)$p.value<.001, .001, round(cor.test(df$IQ, df$EQ)$p.value, 3))
r_value <- round(cor.test(df$IQ, df$EQ)$estimate, 2)
r_square <- round((cor.test(df$IQ, df$EQ)$estimate)*(cor.test(df$IQ, df$EQ)$estimate), 2)

text_results <- paste("The results of the study suggest that IQ ", ifelse(p_value<.05, "is", "is not"), " associated with EQ, r = ",r_value, ", p = ", p_value, ", R^2 = ", r_square,".", sep="")

```


```{r results="asis", eval=cor_ans=="Yes"}
cat(text_results)
```

\newpage

## Formulas: 

Mean = $\frac{\sum{x}}{n}$

Population variance = $\sigma^2 = \frac{\displaystyle\sum_{i=1}^{n}(x_i - \mu)^2} {n}$

Sample variance = $var = \frac{\displaystyle\sum_{i=1}^{n}(x_i - \mu)^2} {n-1}=\frac{SS_x} {n-1}$


Sample standard deviation = $SD = \sqrt{\frac{\displaystyle\sum_{i=1}^{n}(x_i - \mu)^2} {n-1}}$

Sum of the products = $SP = \sqrt{\displaystyle\sum_{i=1}^{n}(x_i - \mu_x)*(y_i - \mu_y)}$

Correlation = $r = \frac{SP}{\sqrt{ss_{x}ss_{y}}}$
